---
title: "COVID-19 Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: cerulean
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(dplyr)
library(plotly)
source("helpers.R")
```

```{r data}

borders = rjson::fromJSON(file="https://raw.githubusercontent.com/johan/world.geo.json/master/countries.geo.json")
country_codes = read.csv("https://raw.githubusercontent.com/hoganj15/MMA_Assignment_Data/main/COVID/country_codes.csv") %>% rename(Country = name, Code = code, lat = latitude, lon = longitude, two_digit = country)
map_data = countries %>% group_by(Country) %>% filter(Date == max(Date))
map_data = left_join(countries, country_codes, by = "Country")
map_data = left_join(map_data, demographics, by = "Country")
map_data$susceptible = map_data$N - map_data$Total
map_data$infected = map_data$Total - map_data$Recovered - map_data$Deaths #active cases are all cases minus ones that have recovered or died
map_data$removed = map_data$Recovered + map_data$Deaths

```


World Active Cases as Percentage of Population
=======================================================================

Column {.sidebar data-width=200}
-------------------------------------------------------------------

The heatmap to the right shows the prevalence of COVID-19 in each country today. Prevalence is calculated as the % of a country's population that are currently infected with COVID-19.


Column {}
-----------------------------------------------------------------------

### Heatmap


```{r}
renderPlotly({
  plot_ly(type = "choroplethmapbox", geojson = borders, locations = map_data$Code, z = 100*map_data$infected/map_data$N, colorscale="Cividis", zmin = 0, zmax = 8, text = ~paste("<br>", "Active Cases:", map_data$infected, "<br>", "Total Cases:", map_data$Total, "<br>", "Total Deaths:", map_data$Deaths), marker = list(line = list(width=0), opacity=0.75)) %>% 
    layout(mapbox = list(style = "open-street-map", zoom = 2.1), title = paste("Active Cases as % of Population"))
})
```


Specific Country Metrics
=======================================================================

Column {.sidebar data-width=200}
-------------------------------------------------------------------
```{r}
selectInput(inputId = "country_chosen", label = "Select Countries:", choices = unique(countries$Country), multiple = TRUE, selected = c("US", "United Kingdom", "Canada", "South Africa", "Netherlands"))
```


Column {data-width=400}
-----------------------------------------------------------------------

### Point Prevalence

Point Prevalence is a metric to capture the proportion of a given country's population that had COVID-19 at a given time. Higher numbers here indicate a period where COVID was very common in that country.


```{r}
renderPlotly({
  fig = plot_ly()
  for (country_chosen in input$country_chosen){fig = fig %>% add_trace(data = plotting_metrics(country_chosen), x = ~day, y = ~100*point_prevalence, mode = 'lines', name = country_chosen)}
  fig = fig %>% layout(title = paste("Point Prevalence"), hovermode = 'x unified', xaxis = list(title = "Day"), yaxis = list(title = "% of Population"))
})
```

Column {data-width=400}
-----------------------------------------------------------------------

### Case Fatality Rate

Case fatality is a measure that captures the proportion of COVID-19 cases that ended up being fatal. High numbers here indicate that a high percentage of people that got COVID-19 were dying in that country and at that time.

```{r}
renderPlotly({
  fig = plot_ly()
  for (country_chosen in input$country_chosen){fig = fig %>% add_trace(data = plotting_metrics(country_chosen), x = ~day, y = ~case_fatality, mode = 'lines', name = country_chosen)}
  fig = fig %>% layout(title = paste("Case Fatality"), hovermode = 'x unified', xaxis = list(title = "Day"), yaxis = list(title = "Fatality Rate"))
})
```


### Cases/Population

This shows the total number of cases over time, as a proportion of the population. Higher numbers here indicate that a higher percentage of the population has gotten COVID-19.

```{r}
renderPlotly({
  fig = plot_ly()
  for (country_chosen in input$country_chosen){fig = fig %>% add_trace(data = plotting_metrics(country_chosen), x = ~day, y = ~Total/N, mode = 'lines', name = country_chosen)}
  fig = fig %>% layout(title = paste("Total Cases as Percentage of Population"), hovermode = 'x unified', xaxis = list(title = "Day"), yaxis = list(title = "$ of Population"))
})
```


